\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2018-02-09 redaction/dhp_july2017.tex
Changement rédaction

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}. Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms. However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much. In this paper, we investigate the robustness of this result. The major contribution of our paper is then to show that it is not robust to a more flexible treatment of the additive component of transport costs (i.e., a cost per unit traded). In contrast to \cite{hummels2007}, we thus find that composition effects play a minor role in accounting for the trend patterns of transport costs, once we allow for a modelling of the additive cost component that varies over country partner and product. More than trade composition effects, it is the reduction of the transport costs ``per se'' that overwhelmingly explains the downward trend of overall transport costs observed over the period.\medskip

@@

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level export data in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price. Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, if our study covers a narrower set of trade costs focused on transport costs, our empirical analysis allows us to provide distinct measures of the levels of both the iceberg and the additive trade costs. We thus obtain that, on average, the additive cost amounts to 1.8\% and 2.9\% of the export price in air and vessel transport respectively; the average estimated value for the ad-valorem component being of 3.5\% and 3.2\% for air/vessel respectively. Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

\textbf{VV: - Il est difficile, de notre propre aveu, de ne pas reconnaitre qu’il y a quand mime de la variation between sectors au niveau 5-digits. Or, nous riagrigeons au niveau 3 ou 4d a cause des NLLS. Pourquoi ne pas estimer l’aquation initiale (juste avant la (5)) en niveau avec des OLS ? A cet agard, l’explication pour la structure multiplicative des rasidus n’est ni tras claire, ni tras convaincante.}

@@

One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport cost burden. Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price). However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, as that would imply some a direct influence between the firm's prices and the size of transport costs, except in the choice between air and vessel transport (that is beyond our scope of analysis). Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control. In this respect, we are confident that our estimation strategy is immune from endogeneity problems. \smallskip

@@

Do additive costs matter? This section provides two elements of answer to this question. First, we quantify the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})). This allows us to estimate whether additive costs represent a sizeable component of transport costs. Second, we assess the importance of the additive component of in overall transport costs through goodness-of-fit measures.

@@

Our first contribution to the literature is to provide estimates for the size of both the ad-valorem and the additive components of transport costs. Table \ref{tab:summary_results} reports a summary of our results. It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data. The first panel reports the results for a specification based on ad-valorem costs estimated alone (Equation (\ref{eq:model_nlI})), while the second panel presents estimates for a specification involving both ad-valorem and multiplicative components (Equation (\ref{eq:model_IetA})). Finally, the third panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level. Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for Air at the 3-digit level removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year. Overall results (over the whole period) are not substantially affected if this year is included though. This can be uncovered from the results detailed year by year in the Online Appendix, available on the authors' webpages.}

@@

\textbf{VV: Nos resultats sont en ad-valorem (multiplicatif du prix), ce qui explique qu’on a des pourcentages superieurs pour Vessel par rapport a Air. Spontanement, on s’attend au contraire (les couts de fret sont plus importants dans l’aerien), mais en fait ces chiffres refletent simplement que les prix sont plus eleves dans l’aerien. En un mot, ces chiffres ne signifient pas que les couts monétaires dans l’aerien sont inférieurs au vessel. Ceci devrait etre précise.}

@@

Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs. Our methodology indeed allows us to go deeper into the structure of transport costs, that can be decomposed into an additive and an ad-valorem components. We thus find that the per-kg cost dimension is sizeable. On average over the whole period, additive costs are estimated to 2.9\% of the export price for ocean shipping, and 1.8\%. This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\/ and 1.8\% for vessel and air shipping respectively (on average over the period). Put it differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward. In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2. These elements point out that additive costs are quantitatively sizeable.

We can use the time dimension of our dataset in order to assess the validity of this result on a year-to-year basis over the period. To this aim, Figure \ref{fig:decomp_TC_3d} reports the shares of both the ad-valorem costs (grey area) and the additive costs (black area) in total costs (the total of the two, expressed in percentage of the fas price), by transport mode for each year between 1974 and 2013.

@@

%\textbf{je ne comprends pas les chiffres mis dans le commentaire en 1974, par rapport aux graphiques.}
%JH: Voir Tableaux 4 et 5, années et modes de transport correspondants. Ils correspondent à la mesure "Iceberg alone", parce que cela semblait permettre une comparaison plus directe avec Hummels. Bref, à discuter.

As reported on Figure \ref{fig:decomp_TC_3d}, total transport costs have regularly declined over time, by an order of magnitude of 3 percentage points in air shipping and 5 percentage points in ocean shipping (basically, divided by 2 in both cases). At this point though, we postpone this discussion about the downward trend for transport costs for Section \ref{sec:results_trends}, to focus on the importance of the additive component. In both transport modes, the additive component appears of sizeable importance all over the period, as the share of additive costs remains roughly constant over time. These results thus suggest that additive costs are neither negligible in magnitude nor an erratic phenomenon. By contrast, they represent an sizeable and structural dimension of international transport costs. Further, one side result of our work is to provide quantitative estimates for the values of both additive and ad-valorem transport costs,which can be very useful for the related more theoretical papers, which need to calibrate their models.\medskip

Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question. Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that the ratio of additive to ad-valorem trade costs amounts to 14\% of the export price, ie $\frac{t/\tau}{\widetilde{p}}=14\%$ expressed in our terminology. One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem). From this, we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect. Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.8\% in air transport. This may sound surprisingly low, in contrast to the 14\% obtained by \citet{Irrazabal_2015}. This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case. While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs (as we start from the gap between the import and the export prices). From \citet{anderson_wincoop_jel}, we can infer that monetary transport costs approximatively represent 15\% of total trade costs.\footnote{\citet{anderson_wincoop_jel} estimate that monetary transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.} Plugging this into \citet{Irrazabal_2015} results, allows to deduce a ratio $t/\tau$ of international transport costs equal to 2.1\% ($0.15\times 0.14$), which is in line with our own results. This makes us confident in our quantitative assessment of the size of the additive component in international transport costs. In the next section, we analyze their importance through another angle based on quality-of-fit diagnostic tests.

\textbf{VV: Les comparaisons chiffrees sont parfois délicates. Le paragraphe de comparaison avec Irarrazabal and co page 12 est tres clair, mais cela vient un peu tard. Il faut que des l’intro on comprenne bien la difference entre notre mesure de coûts additifs (representant entre 40 et 60\% des coûts de transport totaux) et les 14\% d’Irrara and co, qui sont un pourcentage du prix fab.}

\subsection{Assessing the importance of additive transport costs}

\section{Transport Costs time trends: The role of the additive component}\label{sec:results_trends}
\subsection{Estimation strategy}
In this section, we investigate the role of the additive component in total transport costs in a historical perspective.
As a first step, we come back to Figure \ref{fig:decomp_TC_3d}, that displays the respective shares of additive and iceberg components in transport costs over time by transport mode. Considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport. \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates. In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:decomp_TC_3d} springs from high oil-shock related transport costs in 1974. However, computing the time trends from 1980 does not dramatically change the picture. The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs. We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} On US data, \citet{Hummels_1999} obtains that overall transport cost declined from 6 \% to 4 \% of the import value between 1974 and 1996. For the same years, we obtain a total decrease from 6.9 to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping. Overall, our results display trends that are close to those reported by \citet{Hummels_1999}, with a magnitude of transport costs that appears higher in ocean shipping than for air transport over the period.

@@

If we share this view with \cite{hummels2007}, we adopt a different strategy for eliminating composition effects, which builds upon our first-stage results.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} Estimation driven in Section \ref{sec:results_decomposition} indeed provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country. Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect (see details below). For each cost component (additive and ad-valorem), comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the ``ceteris paribus'' costs (for instance, insurance or handling costs) that have reduced over time. We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).}

@@

>>>>

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables}. Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms. However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much. In this paper, we investigate the robustness of this result. As major contribution of the paper, we find that composition effects play a minor role in accounting for the trend patterns of transport costs, in contrast to \cite{hummels2007}. Importantly, we show that this difference of result can be accounted for by the modelling of the additive component of international transport costs (i.e., a transport cost per unit traded). Once we allow for a more flexible (and empirically relevant) modelling of the additive cost component that varies over country partner and product, it is the reduction of the transport costs ``per se'' that overwhelmingly explains the downward trend of overall transport costs observed over the period, rather than trade composition effects. \medskip

@@

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level export data in 2004, they find that the ratio of additive to ad-valorem trade costs is about 14\% of the export price. Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, if our study covers a narrower set of trade costs focused on transport costs, our empirical analysis allows us to provide distinct measures of the levels of both the iceberg and the additive trade costs. We thus obtain that, on average, the additive cost amounts to 1.8\% and 2.9\% of the export price in air and vessel transport respectively; the average estimated value for the ad-valorem component being of 3.5\% and 3.2\% for air/vessel respectively. From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect. Straightforward calculation yields a ratio of additive to multiplicative international transport costs around 2\% (expressed in terms of the export price) in both \citet{Irrazabal_2015} and our paper.\footnote{See Section \ref{sec:results_decomposition} for more details.} This makes us confident in our quantitative assessment of the sizes of international transport costs. Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} or \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

\textbf{VV: - Il est difficile, de notre propre aveu, de ne pas reconnaitre qu il y a quand meme de la variation between sectors au niveau 5-digits. Or, nous reagrigeons au niveau 3 ou 4d a cause des NLLS. Pourquoi ne pas estimer l'equation initiale (juste avant la (5)) en niveau avec des OLS ? A cet egard, l'explication pour la structure multiplicative des residus n'est ni tras claire, ni tres convaincante.}

@@

One may be concerned that the specification of Equations (\ref{eq:model_IetA}) or (\ref{eq:model_nlI}) might be subject to endogeneity bias, as the price set be the exporter may vary depending on the transport cost burden. Studies on the pricing-to-market behavior of firms (see \citealp{Krugman-87}) show that we cannot exclude that the export price set by the firm ($\widetilde{p}_{ik}$) is partly endogenous to the size of transport costs (for instance, the exporting firm absorbing (part of) the transport costs by reducing the fas price). However, we consider as a reasonable assumption that the exporting firm cannot influence the international transport cost sector, that would imply a causal effect of the firm's prices on the transport costs value, except in the choice between air and vessel transport (that is beyond our scope of analysis). Otherwise stated, conditional to an export price and a transport mode, the gap between the export price declared to the custom data and the import price recorded by the US administration is beyond the firm's control. In this respect, we are confident that our estimation strategy is immune from endogeneity problems. \smallskip

@@

Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components. The first step of analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})). This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizeable component of transport costs. In this respect, these results \textit{per se} also constitute our first original contribution to the literature. In Appendix \ref{app:diagnostic_test}, we provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation.

@@

Table \ref{tab:summary_results} reports a summary of our results. It displays the mean and median values of each type of transport costs (either ad-valorem estimated alone or estimated along with additive costs), as well as the associated standard deviation, averaged over the period 1974-2013, for estimation driven both at the 3- and 4-digit sectorial level for different specifications and data. The first panel reports the results for a specification based on ad-valorem costs estimated alone (Equation (\ref{eq:model_nlI})), while the second panel presents estimates for a specification involving both ad-valorem and multiplicative components (Equation (\ref{eq:model_IetA})). Finally, the third panel reports the same set of descriptive statistics, but for the actual cif/fas ratio in our data.\footnote{In Appendix \ref{app:more_results}, we report similar results for a sample of years, for both transport mode, at the 3 and 4-digit classification level. Results for all years (available at the 3-digit level) are reported in the Online Appendix, available on the authors' webpages.}$^{,}$\footnote{We present the estimation results for Air at the 3-digit level removing the year 1989, as the results reveal the presence of strong outliers that bias the estimates of transport costs upwards this particular year. Overall results (over the whole period) are not substantially affected if this year is included though. This can be uncovered from the results detailed year by year in the Online Appendix, available on the authors' webpages.}

@@

\textbf{VV: Nos resultats sont en ad-valorem (multiplicatif du prix), ce qui explique qu’on a des pourcentages superieurs pour Vessel par rapport a Air. Spontanement, on s’attend au contraire (les couts de fret sont plus importants dans l’aerien), mais en fait ces chiffres refletent simplement que les prix sont plus eleves dans l’aerien. En un mot, ces chiffres ne signifient pas que les couts monétaires dans l’aerien sont inférieurs au vessel. Ceci devrait etre precise.}

\textbf{Reponse de Lise: Une facon de repondre a ca est de montrer le prix fob moyen dans l'aerien / vessel. Pour montrer que c'est plus eleve dans l'aerien. Fournir une estimation en \$ du cout par unit (kg) additif pourrait etre interessant aussi, apres tout c'est fondamentalement ca qu'ils sont.}

@@

Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs. Our methodology indeed allows us to go deeper into the structure of transport costs, that can be decomposed into an additive and an ad-valorem components. We thus find that the per-kg cost dimension is quantitatively sizeable. On average over the whole period, additive costs are estimated to 2.9\% of the export price for ocean shipping, and 1.8\%. This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 1.8\% for vessel and air shipping respectively (on average over the period). Put it differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward. In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2. Further, these results provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for the related more theoretical papers, which need to calibrate their models.\smallskip

Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question. Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that the ratio of additive to ad-valorem trade costs amounts to 14\% of the export price, ie $\frac{t/\tau}{\widetilde{p}}=14\%$ expressed in our terminology. One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem). From this, we can rebuilt the ratio in similar terms to them -- allowing us to gain in generality in this respect. Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.8\% in air transport. This may sound surprisingly low, in contrast to the 14\% obtained by \citet{Irrazabal_2015}. This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case. While the database of \citet{Irrazabal_2015} allows to study trade costs in general, our database covers a subset made of the monetary international transport costs (as we start from the gap between the import and the export prices). From \citet{anderson_wincoop_jel}, we can infer that monetary transport costs approximatively represent 15\% of total trade costs.\footnote{\citet{anderson_wincoop_jel} estimate that monetary transport costs represent a 11\% markup over production costs, and that total international trade costs represent 74\%, so transport costs amount to 11/74= 15\% of total trade costs.} Plugging this into \citet{Irrazabal_2015} results, allows to deduce a ratio $t/\tau$ of international transport costs to the export price $\widetilde{p}$ equal to 2.1\% ($0.15\times 0.14$). This stands in line with our own results, and makes us confident in our quantitative assessment of the size of the additive component in international transport costs.\footnote{See also Appendix \ref{app:diagnostic_test}, which confirms the importance of the additive cost component through another perspective angle based on quality-of-fit diagnostic tests.}\smallskip

\medskip

\textbf{VV: Les comparaisons chiffrees sont parfois délicates. Le paragraphe de comparaison avec Irarrazabal and co page 12 est tres clair, mais cela vient un peu tard. Il faut que des l’intro on comprenne bien la difference entre notre mesure de coûts additifs (representant entre 40 et 60\% des coûts de transport totaux) et les 14\% d’Irrara and co, qui sont un pourcentage du prix fab. MIS UN TOPO LA DESSUS EN INTRO. Ok AINSI?}

\section{Transport Costs time trends: The role of the additive component}\label{sec:results_trends}

In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database. This focus is shared with part of the international trade literature, such as \cite{hummels2007} and \cite{Behar_Venables}. Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). \cite{Glaeser04} find that, over the twentieth century, the cost of moving goods have declined by over 90\% in real terms. However, \cite{hummels2007} points out that trade composition effects dominate the characterization of time trends of international transport costs and that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods and origins) have not declined much. In this paper, we investigate the robustness of this result.

\subsection{A first look at the trend patterns}

As a first step, Figure \ref{fig:decomp_TC_3d} displays the respective shares of additive (black area) and iceberg (grey area) components in total transport costs in total costs (the total of the two, expressed in percentage of the fas price), by transport mode for each year between 1974 and 2013.

@@

Two main comments can be made. First, in both transport modes, the additive component appears of sizeable importance all over the period, as the share of additive costs remains roughly constant over time. Confirming the results obtained in Section \ref{sec:results_decomposition}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon. By contrast, they represent an sizeable and structural dimension of international transport costs.

Second, considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport. \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates. In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:decomp_TC_3d} springs from high oil-shock related transport costs in 1974. However, computing the time trends from 1980 does not dramatically change the picture. The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs. We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} On US data, \citet{Hummels_1999} obtains that overall transport cost declined from 6 \% to 4 \% of the import value between 1974 and 1996. For the same years, we obtain a total decrease from 6.9 to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping. Overall, our results display trends that are close to those reported by \citet{Hummels_1999}, with a magnitude of transport costs that appears higher in ocean shipping than for air transport over the period.\smallskip

@@

\paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}, that starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f$ the shipping charge per kg shipped. Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p_{ikt}}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, the transport cost measure writes down as:

\begin{equation}
TC_{ikt}\equiv \frac{p_{ikt}-\widetilde{p}_{ikt}}{\widetilde{p}_{ikt}} = \widetilde{p}_{ikt}^{-\beta}X_{ikt} \label{eq:Hummel_benchmarkeq}
\end{equation}

Notice that $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value). As such, a value of $\beta = 0$ corresponds to the standard case of ``iceberg'' costs, in which transport costs are purely ad valorem. At the other extreme, $\beta = 1$ means that all transport costs are additive. Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs. Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which Hummels (2007) decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described with more details in Appendix \ref{app:compare_Hummels}. At this point though, let us mention one key assumption retained by Hummels (2007), which assumes $\beta$ constant across the triplet origin country origin/product/year. In light of the economic interpretation of $\beta$, this means that the share of additive costs in total costs does not vary across these dimensions.

However, in light of our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and Figure \ref{fig:decomp_TC_3d}, one may wonder about the empirical relevance of this assumption. As reported in Figure \ref{fig:decomp_TC_3d}, it seems that if overall transport costs have decreased over time, its is also the case for each additive and multiplicative components. Further, one may wonder about the constancy of the share of additive costs along the couple (product/origin country).

We investigate this point deeper by...

\textbf{stop here, A FINIR}

\subsection{Estimation strategy}

The above descriptive statistics analysis drives us to adopt a different strategy from \cite{hummels2007} for eliminating composition effects, which builds upon our first-stage results.\footnote{See Appendix \ref{app:compare_Hummels} for a detailed comparison of Hummel's (2007) methodology and ours.} Estimation driven in Section \ref{sec:results_decomposition} indeed provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country. Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect (see details below). For each cost component (additive and ad-valorem), comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the ``ceteris paribus'' costs (for instance, insurance or handling costs) that have reduced over time. We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).}

@@

\subsection{Assessing the importance of additive transport costs \label{app:diagnostic_test}}
